Aligning marine species range data to better serve science and conservation

### 2016-12-30 data_explore/data_explore.Rmd
updated the figures and checked the new reference numbers...

taxa_df <- read_csv(file.path(dir_data, 'taxa_df.csv'))

@@

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps) + sum(Paired),
            IUCN_total = sum(IUCN) + sum(Paired),
            n_total  = sum(n_total),
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Paired_end = AquaMaps_start,
            Paired_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Paired) %>%
              arrange(desc(n_total))) %>%
  bind_rows(x <- taxa_omitted %>%
              group_by(class) %>%
              summarize(n_total = sum(n_am_class_all),
                        dataset = 'AquaMaps',
                        start = 0,
                        end = 100)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
         dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))

>>>>

taxa_df <- read_csv(file.path(dir_data, 'taxa_df.csv'))
taxa_omitted <- read_csv(file.path(dir_data, 'taxa_omitted.csv'))

n_spp_omitted <- taxa_omitted$n_am_class_all %>% sum()

@@

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps)  + n_spp_omitted + sum(Paired),
            IUCN_total = sum(IUCN) + sum(Paired),
            n_total  = sum(n_total) + n_spp_omitted,
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Paired_end = AquaMaps_start,
            Paired_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Paired) %>%
              arrange(desc(n_total))) %>%
  bind_rows(taxa_omitted %>%
              group_by(class) %>%
              summarize(n_total = sum(n_am_class_all),
                        dataset = 'AquaMaps',
                        start = 0,
                        end = 100)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
         dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))

### 2016-12-30 manuscript_main.Rmd
updated the figures and checked the new reference numbers...

<!--
_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
-->

@@

<!--
| Figure  |  w x h (px) | w x h (cm)  | char # |
| :------ |     ------: |     ------: |------: |
| 1 A - C | 2102 x 1536 | 17.8 x 13.0 |   4680 |
| caption |             |             |    367 | 
| 2 A - B | 2102 x 1940 | 17.8 x 16.5 |   5940 |
| caption |             |             |    934 | 
|       3 | 1027 x  531 |  8.7 x  4.5 |    810 |
| caption |             |             |    744 | 
|       4 | 1346 x  590 | 11.4 x  5.0 |   1800 |
| caption |             |             |    870 | 
|  total  |             |             |  16145 |
-->

>>>>

<!--
_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, area of occupancy, spatial ecology
-->

@@

### 2016-12-30 ms_3_res_disc_concl.Rmd
updated the figures and checked the new reference numbers...

Because corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we explored implications of explicitly restricting IUCN ranges to depths based on species' life histories.  This adjustment was not necessary for AquaMaps data because models explicitly include ocean depth preference as a parameter.  While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [REF IS THIS IN THE TRAINING MATERIALS? @iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted. Additionally, IUCN Red List mapping standards formerly required, and still allow, a 50 km buffer around the coastline for coastal species [iucn_iucn_2015-1]; such a buffer directly conflicts with habitat limitations (such as depth) and distorts our understanding of species distribution.

>>>>

Because corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we explored implications of explicitly restricting IUCN ranges to depths based on species' life histories.  This adjustment was not necessary for AquaMaps data because models explicitly include ocean depth preference as a parameter.  While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." REF IS THIS IN THE TRAINING MATERIALS? [@iucn_iucn_2015]), it is not presented as a requirement, so we cannot take its inclusion for granted. Additionally, IUCN Red List mapping standards formerly required, and still allow, a 50 km buffer around the coastline for coastal species [@iucn_red_2015]; such a buffer directly conflicts with habitat limitations (such as depth) and distorts our understanding of species distribution.
